---
title: "Body size-Data analysis"
author: "Carlos"
date: "Tuesday, March 17, 2015"
output: html_document
---
```{r}
library(knitr)
opts_chunk$set(warning = FALSE)
```

#Read data

```{r}
#Read data
#bodysizes<-read.delim('bodysizes_2008.txt')
#explore structure
#str(bodysizes)

```

##subset a particular dataset

Subset the food web for the benguela ecosystem and save it for future use
```{r}
#benguela <- subset(bodysizes,Geographic.location == "Africa, Benguela ecosystem")
#write.csv(benguela,'benguela.csv',row.names = FALSE)
```

```{r}
benguela<-read.csv('benguela.csv',na.string=-999)
```

##Format data for inputing into Igraph

```{r}
# Put the columns in icols to the start of the dataframe
ReorderColumns<-function(DF,icols){
  cols <- c(icols,names(DF)[-which(names(DF) %in% icols)])
  DF <- DF[cols]
  return(DF)
}

icols <- c("Common.name.s..resource","Common.name.s..consumer")

benguela<-ReorderColumns(benguela,icols)

```

```{r}
library(igraph)
#create graph
G <- graph.data.frame(benguela)
plot(G)
```

##Finding IGP modules

```{r}

findIGPs<-function(G){
  FocusVertices <- V(G)[degree(G,mode="out") >=2]
  IGPs <- findModules(FocusVertices,G)
  return(IGPs)
}

findModules<-function(V,G){
  Modules <- c()
  for(resource in V){
    Modules<-append(Modules,IGPmod(G,resource))
  }
  return(Modules)
}  
IGPmod<-function(G,resource){
  Predators <- neighborhood(G,1,nodes = resource,mode ="out")
  IGPs <- c()
  n <- length(Predators[[1]])
  for(i in 2:n){
    IGPs<-append(IGPs,findIGPrey(resource,i,Predators,G))
  }
  return(IGPs)
}

findIGPrey<-function(resource,predindex,Predators,G){
  pIGPs<-c()
  NeighPred <- neighborhood(G,1, nodes = Predators[[1]][predindex],mode="out")
  for(p in Predators[[1]][-c(1,predindex)]){
    if(p %in% NeighPred[[1]]){
      pIGPs<-append(pIGPs,list(c(resource,Predators[[1]][predindex],p)))
    }
  }
  return(pIGPs)
}
```

```{r}
IGPs<-findIGPs(G)
```

##Finding body size ratios within IGP modules
For the three species of an IGP module , *R,C,P* we will find :  
$$
\begin{equation}
\begin{aligned}
k_{RC} := \frac{m_R}{m_C} \\
k_{CP} := \frac{m_C}{m_P} 
\end{aligned}
\end{equation}
$$
```{r}

##Finding size ratios for the IGP modules

findIGPsizeRatios <- function(IGPs,G){
  log10.K_RC<-c()
  log10.K_CP<-c()
  log10.m_P <-c()
  for(i in 1:length(IGPs)){
    SR <- getsizeRatios(G,IGPs[i][[1]])
    log10.K_RC[i] = log10(SR[1])
    log10.K_CP[i] = log10(SR[2])
    log10.m_P[i] = log10(SR[3])
  }
  
  return(data.frame(log10.K_RC,log10.K_CP,log10.m_P))
}

getsizeRatios<-function(G,igpL){
  SR<-c()
  igpEdges = getIGPedges(G,igpL)
  for(edge in igpEdges){
    E <- E(G)[edge]
    
    SR<-append(SR,c(1/E$Consumer.resource.body.mass.ratio))
  }
  
  m_P =E(G)[igpEdges[2]]$Mean.mass..g..consumer
  SR[3] <- m_P
  return(SR)
  
}

getIGPedges<-function(G,igpL){
  edgeInd<-c()
  for(i in 1:2){
    edgeInd<-append(edgeInd,c(igpL[i],igpL[i+1]))
  }
  edges<-get.edge.ids(G,edgeInd)
  return(edges)
}
```


```{r}
SR <- findIGPsizeRatios(IGPs,G)

SR$m_pInt <- as.factor(floor(SR$log10.m_P))
```


##Exploratory Plots

code to plot regression equation taken from <http://stackoverflow.com/questions/7549694/ggplot2-adding-regression-line-equation-and-r2-on-graph>

```{r}
lm_eqn = function(m) {

  l <- list(a = format(coef(m)[1], digits = 2),
      b = format(abs(coef(m)[2]), digits = 2),
      r2 = format(summary(m)$r.squared, digits = 3));

  if (coef(m)[2] >= 0)  {
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,l)
  } else {
    eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)^2~"="~r2,l)    
  }

  as.character(as.expression(eq));                 
}

```

```{r}

library(ggplot2)
reg1<-lm(log10.K_RC~log10.K_CP , data = SR)
summary(reg1)
plot <- ggplot(SR,aes(log10.K_CP,log10.K_RC)) + geom_point(aes(color=m_pInt))+labs(x = expression("log "* K[CP]), y = expression("log "* K[RC])) + geom_smooth(method="lm",col="red") +theme_bw(base_size = 10) 

plot + annotate("text", x = -6 , y = -5, label = lm_eqn(reg1), colour="black", size = 5, parse=TRUE)

```

#Automatizing all the job

```{r}
getIGPSizeRatios <- function(Data,Location,icols,Shabitat = FALSE){
  Subdata <- Data[Data$Geographic.location == Location,]
  if(class(Shabitat) != "logical"){
    Subdata<-Subdata[Subdata$Specific.habitat == Shabitat,]
  }
  Subdata <- ReorderColumns(Subdata,icols)
  G_Subdata <- graph.data.frame(Subdata)
  IGPs <- findIGPs(G_Subdata)
  SR <- findIGPsizeRatios(IGPs,G_Subdata)
  
  SR$m_pInt <- as.factor(floor(SR$log10.m_P))
  
  reg1<-lm(log10.K_RC~log10.K_CP , data = SR)
  print(summary(reg1))
  
  ##Plotting
  plot <- ggplot(SR[complete.cases(SR$log10.m_P),],aes(log10.K_CP,log10.K_RC)) + geom_point(aes(col = m_pInt))+
          labs(x = expression("log "* K[CP] ), y = expression("log "* K[RC])) + 
          geom_smooth(method="lm",col="red") +theme_bw(base_size = 10)+
          annotate("text", x = min(SR$log10.K_CP) + 1 , y = min(SR$log10.K_RC)+1,
                   label = lm_eqn(reg1), colour="black", size = 3, parse=TRUE)
  print(plot)
  return(SR)
}

```

#Silkwood Park food web
```{r , warning=FALSE}
bodysizes<-read.csv('bodysizes.csv',na.string = -999)
bodysizes<-bodysizes[complete.cases(bodysizes$Geographic.location),]

Silwood<-getIGPSizeRatios(bodysizes,"Country: United Kingdom; UTM: 51.24'N, 0.34'W; Silwood Park, Berkshire",c("Taxonomy.resource","Taxonomy.consumer"))
```
  
#Tuesday lake

```{r}
TL<-getIGPSizeRatios(bodysizes,"Country: United States of America; UTM: 89.32'W, 46.13'N; Tuesday Lake, Michigan ",c("Taxonomy.resource","Taxonomy.consumer"))
```

#Broadstone stream

```{r}
Broadstone<-getIGPSizeRatios(bodysizes,"Country: United Kingdom; UTM: 51.05'N, 0.03'E; Broadstone Stream in Sussex",c("Taxonomy.resource","Taxonomy.consumer"))
```

#NorthYorkshire

```{r}
NorthYorkshire<-getIGPSizeRatios(bodysizes,"Country: United Kingdom; UTM: 53.40'N, 0.59'W; Skipwith Common, North Yorkshire",c("Taxonomy.resource","Taxonomy.consumer"))
```

#Nevada Mountains

```{r}
Nevada<-getIGPSizeRatios(bodysizes,"Country: United States of America; John Muir Wilderness Area and Kings Canyon National Park, Sierra Nevada mountains, California",c("Taxonomy.resource","Taxonomy.consumer"))
```

#Weddell Sea

```{r}
Weddell<-getIGPSizeRatios(bodysizes,"Antarctica: Eastern Weddell Sea",c("Taxonomy.resource","Taxonomy.consumer"))
```

#Mill Stream UK

```{r}
Mill<-getIGPSizeRatios(bodysizes,"Country: United Kingdom; the River Frome, Dorset;  UTM: 50.40'N, 2.11'W",c("Taxonomy.resource","Taxonomy.consumer"))
```

#Parasitoid community

```{r}
ParaC<-getIGPSizeRatios(bodysizes,"Country: United Kingdom",c("Taxonomy.resource","Taxonomy.consumer"))
```

#Grand Caricaie 

##Grand caricaie mown marsh dominated by Cladetium

```{r}
GrandCaricaie1<-getIGPSizeRatios(bodysizes,"Country: Switzerland; Lake Neuchâtel",c("Taxonomy.resource","Taxonomy.consumer"),Shabitat = "Grand Caricaie; marsh dominated by Cladietum marisci, mown; Clmown1")

```


##Grand caricaie marsh dominated by Schoneus mown

```{r}
GrandCaricaie2<-getIGPSizeRatios(bodysizes,"Country: Switzerland; Lake Neuchâtel",c("Taxonomy.resource","Taxonomy.consumer"),Shabitat = "Grand Caricaie; marsh dominated by Schoenus nigricans, mown; Scmown2;")
```

#Agregate all data sets

```{r}
total<-rbind(SR,Silwood,TL,Broadstone,NorthYorkshire,Nevada,Weddell,Mill,ParaC,GrandCaricaie1)
```

##Plot total

```{r}
total$m_pInt <- factor(floor(total$log10.m_P))
reg<-lm(log10.K_RC~log10.K_CP,data=total)
summary(reg)
plot <- ggplot(total[complete.cases(total$log10.m_P),],aes(log10.K_CP,log10.K_RC)) + 
        geom_point(aes(col = m_pInt))+
        labs(x = expression("log "* K[CP] ), y = expression("log "* K[RC])) + 
        geom_smooth(method="lm",col="red") +theme_bw(base_size = 10) 

plot + annotate("text", x = min(total$log10.K_CP) + 2 , y = min(total$log10.K_RC)+2, 
                  label = lm_eqn(reg), colour="black", size = 3, parse=TRUE)

```

##Linear model
$$
\log_{10}(k_{RC}^i) \approx a_0 + a_1\log_{10}(k_{CP}^i) + a_2\log_{10}(m_P^i) 
$$

```{r}
reg2 <- lm(log10.K_RC~log10.K_CP+log10.m_P ,data = total[complete.cases(total$log10.m_P),])
summary(reg2)
```



